---
title: "p8105_hw2_hw2849"
author: hw2849
date: 2021.10.04
output: github_document
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Use excel `Mr. Trash Wheel` dataset

```{r, message = FALSE, warning = FALSE}
litter_data = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>% 
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls, 0)
         )

head(litter_data, 5)
```

Clean precipitation data for 2019.

```{r, message = FALSE, warning = FALSE}
prep_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>% 
  mutate(
    year = 2019, month = month.name[as.numeric(month)]
) 

head(prep_2019, 5)
```

Clean precipitation data for 2018.

```{r, message = FALSE, warning = FALSE}
prep_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, "A2:B14") %>% 
   janitor::clean_names() %>%
   drop_na() %>% 
  mutate(
    year = 2018, month = month.name[as.numeric(month)]
) 

head(prep_2018, 5)
```

Combine two datasets

```{r, message = FALSE, warning = FALSE}
combine_data = full_join(prep_2018, prep_2019)

head(combine_data, 5)
```

Find the total Precipitation in 2018.

```{r, message = FALSE, warning = FALSE}
combine_data %>% 
filter(year == "2018") %>% 
  pull(total) %>% 
  sum()
```

Find the median number of sports balls in a dumpster in 2019

```{r, message = FALSE, warning = FALSE}
filter(litter_data, year == "2019") %>% 
  pull(sports_balls) %>% 
  median()
```

`litter_data` dataset contains 344 observations and 17 variables, including years, months, dates, and different kinds of trash such as sports balls. Cleaned precipitation data for 2018 has 12 observations pf 3 variables, while for 2019 is 6 observations and 3 variables. The combined table has 18 observations of 3 variables in total. The total precipitation in 2018 was 70.33, and the median number of sports balls in a dumpster in 2019 was 9. 

## Problem 2

Let's tidy `pols-month.csv`.

```{r, message = FALSE, warning = FALSE}
pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  
## separate `mon` into `ymd` and convert month to character
  separate(mon, c("year", "month", "day")) %>% 
  mutate(
    month = month.name[as.numeric(month)], 
  ) %>% 
  
## create a president variable with values of gop and dem
  pivot_longer(
    prez_gop | prez_dem,
    names_to = "president",
    names_prefix = "prez_",
    values_to = "gop_dem"
  ) %>% 
  filter(gop_dem != 0) %>% 
  
## get rid of the column of day and prez_gop and prez_dem
 select(-day, -gop_dem)

head(pols_month, 5)
```

clean the data in `snp.csv`

```{r, message = FALSE, warning = FALSE}
snp_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  
## Separate `date` into `ymd`
  mutate(
    date = lubridate::mdy(date)
  ) %>% 
  
## Convert month to character and make year varibale true to the reality
  separate(date, c("year", "month", "day")) %>% 
  mutate(
    month = month.name[as.numeric(month)],
   year = as.numeric(year),
   year = ifelse(year > 2021, year - 100, year)
  ) %>% 
  mutate(year = as.character(year)) %>% 

## remove the day variable
  select(-day)

head(snp_df, 5)
```

Tidy the unemployment data

```{r, message = FALSE, warning = FALSE}
unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  rename(year = Year) %>% 
  
## organize this data with pivot_longer
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
    ) %>% 
  mutate(year = as.character(year))

head(unemployment, 5)
```

Join the datasets by merging `snp` into `pols`, and merging `unemployment` into the result.

```{r, message = FALSE, warning = FALSE}
## join snp_df into pols_month table with left_join function
snp_into_pols = left_join(pols_month, snp_df)

## merging unemployment into the table above 
results = left_join(snp_into_pols, unemployment)

head(snp_into_pols, 5)
head(results, 5)
```

In a summary, we used the `FiveThirtyEight` dataset. The `pols-month.csv` table contains 822 rows and 9 variables, including year, month, and president in that year. `snp_df` table contains 787 rows and 3 columns with variables year, month, and close. And `unemployment` table contains 816 rows and 3 columns with variables year, month, and unemployment rate. Then we combined all three tables together by common variable `year` and `month`, getting new tables `results`, which contains 822 rows and 11 variables. 


## Problem 3

Let's tidy the dataset of NYC popularity of baby names. 

```{r, message = FALSE, warning = FALSE}
baby_name = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  
## remove duplicated rows
  distinct() %>% 
  
## organize the names of a categorical predictor 
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
      ethnicity = replace(ethnicity,  ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"), 
      ethnicity = replace(ethnicity,  ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")
  ) %>%

## solve case structure of string variables (child's name)
  mutate(
    childs_first_name = str_to_sentence(childs_first_name)
  )

head(baby_name, 5)

## creating a table showing the rank in popularity of Olivia(female) over time
olivia_df = baby_name %>% 
select(-count) %>% 
  filter( gender == "FEMALE", childs_first_name == "Olivia") %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )

## creating a table for most popular male name over time
most_popular_male_name = baby_name %>% 
  filter(gender == "MALE", rank == "1") %>% 
select(-rank, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )

## a scatterplot for number of children with a name vs rank in popularity
baby_name %>% 
  filter(
    gender == "MALE", 
    year_of_birth == "2016", 
    ethnicity == "WHITE NON HISPANIC"
    ) %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point() + 
  labs(
    title = "Popularity White Non-Hispanic Male Children's names in 2016",
       x = "Rank in popularity of names",
       y = "Number of children with a name"
)
```

In this problem, we removed duplicated rows and then organized the categorical names and names' cases structure, so the `baby_name` table contains 12181 observations with 6 variables, such as year of birth, gender, ethnicity, names, and popularity of the name. Next, we created a table for the female name `Olivia` to see the popularity over years. And we did similar process to the most popular male names over time. At last, we drew a scatterplot to show the rank in popularity of names vs the counts of children with a name. 

