---
title: "p8105_hw2_hw2849"
author: hw2849
date: 2021.10.04
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)

```

## Problem 1

Use excel `Mr. Trash Wheel` dataset

```{r}
litter_data = read_excel("./data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls, 0)
         )
litter_data
```

Clean precipitation data for 2019.

```{r}
prep_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = 7, "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>% 
  mutate(
    year = 2019, month = month.name[as.numeric(month)]
) 
```

Clean precipitation data for 2018.

```{r}
prep_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = 8, "A2:B14") %>% 
   janitor::clean_names() %>%
   drop_na() %>% 
  mutate(
    year = 2018, month = month.name[as.numeric(month)]
) 
```

Combine two datasets

```{r}
combine_data = full_join(prep_2018, prep_2019)
```

Find the total Precipitation in 2018.

```{r}
combine_data %>% 
filter(year == "2018") %>% 
  pull(total) %>% 
  sum()
```

Find the median number of sports balls in a dumpster in 2019

```{r}
filter(litter_data, year == "2019") %>% 
  pull(sports_balls) %>% 
  median()
```

`litter_data` dataset contains 344 observations and 17 variables, including years, months, dates, and different kinds of trash such as sports balls. Cleaned precipitation data for 2018 has 12 observations pf 3 variables, while for 2019 is 6 observations and 3 variables. The combined table has 18 observations of 3 variables in total. The total precipitation in 2018 was 70.33, and the median number of sports balls in a dumpster in 2019 was 9. 

## Problem 2

Let's tidy `pols-month.csv`.

```{r}
pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  
## separate `mon` into `ymd` and convert month to character
  separate(mon, c("year", "month", "day")) %>% 
  mutate(
    month = month.name[as.numeric(month)], 
  ) %>% 
  
## create a president variable with values of gop and dem
  pivot_longer(
    prez_gop | prez_dem,
    names_to = "president",
    names_prefix = "prez_",
    values_to = "gop_dem"
  ) %>% 
  filter(gop_dem != 0) %>% 
  
## get rid of the column of day and prez_gop and prez_dem
 select(-day, -gop_dem)

```

clean the data in `snp.csv`

```{r}
snp_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  
## Separate `date` into `ymd`
  mutate(
    date = lubridate::mdy(date)
  ) %>% 
  
## Convert month to character and make year varibale true to the reality
  separate(date, c("year", "month", "day")) %>% 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
     year = ifelse(year > 2021, year - 100, year)
  ) %>% 
  mutate(year = as.character(year)) %>% 

## remove the day variable
  select(-day)

```

Tidy the unemployment data

```{r}
unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  rename(year = Year) %>% 
  
## organize this data with pivot_longer
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
    ) %>% 
  mutate(year = as.character(year))
```

Join the datasets by merging `snp` into `pols`, and merging `unemployment` into the result.

```{r}
## join snp_df into pols_month table with left_join function
snp_into_pols = left_join(pols_month, snp_df)

## merging unemployment into the table above 
results = left_join(snp_into_pols, unemployment)
```


## Problem 3

Let's tidy the dataset of NYC popularity of baby names. 

```{r}
baby_name = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  
## remove duplicated rows
  distinct() %>% 
  
## solve case structure of string variables (child's name)
  mutate(
    childs_first_name = str_to_sentence(childs_first_name)
  )

## creating a table showing the rank in popularity of Olivia(female) over time
olivia_df = filter(baby_name, gender == "FEMALE", childs_first_name == "Olivia")

## creating a table for most popular male name over time
most_popular_male_name = filter(baby_name, gender == "MALE", rank == "1")

baby_name %>% 
  filter(
    gender == "MALE", 
    year_of_birth == "2016", 
    ethnicity == "WHITE NON HISPANIC"
    ) %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point() + 
  labs(
    title = "Popularity White Non-Hispanic Male Children's names in 2016",
       x = "Rank in popularity of names",
       y = "Number of children with a name"
)
```

